graphsim: An R package for simulating gene expression data from graph structures of biological pathways

### 2020-05-09 paper/paper.Rmd
correct figure formatting and labels  (fig2-4, pane B-D)

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", width = 68)
knitr::opts_chunk$set(fig.cap = "", fig.path = "Plot")
options(width = 68, cli.unicode = FALSE, cli.width = 68)
par(mai=c(1.02, 0.82, 0.82, 0.82)-0.82)
par(bty="o")
#captioner::captioner(prefix = "Fig.")
```

@@

plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.05, adj=0.5, cex=1.75)
box()

@@

plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(b) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
box()
```

@@

```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}. An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes. Non-negative values are coloured white to red from $0$ to $1$. This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.'}
# activating graph
state <- rep(1, length(E(graph)))
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
mtext(text = "(a) Activating pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
box()

@@

op <- par(mfrow=c(2,2), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
#plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(b) relationship matrix", side=1, line=3, at=0.075, adj=0.5, cex=1.75)

#plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)

@@

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)

op <- par(mfrow=c(1,1), mar=c(4,4,4,1), oma=c(0.5,0.5,0.5,0))
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
par(op)
```

@@

@@

#generate parameters for inhibitions
state <- c(1, 1, -1, 1, 1, 1, 1, -1)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
mtext(text = "(a) Inhibiting pathway structure", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
box()

@@

# #plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(b) relationship matrix", side=1, line=3, at=0.075, adj=0.5, cex=1.75)

# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)

@@

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)

#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
```

@@

# #plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)),
              sub = "(b) relationship matrix", cex.sub = 1.75)

# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3, at=0.075, adj=0.5, cex=1.75)

@@

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(d) Simulated correlation", side=1, line=3, at=0.075, adj=0.5, cex=1.75)

#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3, at=0.075, adj=0.5, cex=1.75)
```

>>>>

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", width = 68)
knitr::opts_chunk$set(fig.cap = "", fig.path = "Plot")
options(width = 68, cli.unicode = FALSE, cli.width = 68)
#par(mai=c(2.82, 2.82, 0.82, 0.82)-0.82)
par(mar=c(7, 10, 4, 2) + 0.1)
par(bty="o")
#captioner::captioner(prefix = "Fig.")
```

@@

plot_directed(graph, state ="activating", layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(a) Activating pathway structure", side=1, line=3.5, at=0.05, adj=0.5, cex=1.75)
box()

@@

plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node = 2, cex.arrow = 4, arrow_clip = 0.2)
mtext(text = "(b) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
box()
```

@@

```{r, include = FALSE}
set.seed(9000)
```

```{r simulation_activating, fig.align='center', fig.show='hold', fig.width='1.0\\linewidth', fig.height='1.0\\linewidth', out.height='.415\\linewidth', out.width='.415\\linewidth', fig.retina=10, fig.margin = FALSE, fig.ncol = 2, fig.cap = '\\textbf{Simulating expression from a graph structure}. An example of a graph structure (a) that has been used to derive a relationship matrix (b), $\\Sigma$  matrix (c) and correlation structure (d) from the relative distances between the nodes. Non-negative values are coloured white to red from $0$ to $1$. This $\\Sigma$ matrix has been used to generate a simulated expression dataset of 100 samples (coloured blue to red from low to high) via sampling from the multivariate normal distribution. Here genes with closer relationships in the pathway structure show a higher correlation between simulated values.'}
# activating graph
state <- rep(1, length(E(graph)))
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
mtext(text = "(a) Activating pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
box()

@@

#plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(b) relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)

#plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)

@@

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)

#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
```

```{r, include=FALSE}
pdf("Plotsimulation_activating-5.pdf", width = 13.44, height = 6.72)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
dev.off()
system("sed -i.bak '/Plotsimulation_activating-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
system("sed -i.bak '/Plotsimulation_activating-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
```

@@

```{r, include = FALSE}
set.seed(9000)
```

@@

#generate parameters for inhibitions
state <- c(1, 1, -1, 1, 1, 1, 1, -1)
plot_directed(graph, state=state, layout = layout.kamada.kawai,
              cex.node=2, cex.arrow=4, arrow_clip = 0.2)
mtext(text = "(a) Inhibiting pathway structure", side=1, line=3.5, at=0.075, adj=0.5, cex=1.75)
box()

@@

#plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(b) relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)

# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)

@@

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)))
mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)

#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
```

```{r, include=FALSE}
pdf("Plotsimulation_inhibiting-5.pdf", width = 13.44, height = 6.72)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
dev.off()
system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
system("sed -i.bak '/Plotsimulation_inhibiting-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
```

@@

# #plot relationship matrix
heatmap.2(make_distance_graph(graph, absolute = FALSE),
          scale = "none", trace = "none", col = colorpanel(50, "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
mtext(text = "(b) relationship matrix", side=1, line=3.5, at=0, adj=0.5, cex=1.75)

# #plot sigma matrix
heatmap.2(make_sigma_mat_dist_graph(graph, state = state, 0.8, absolute = FALSE),
scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
mtext(text = expression(paste("(c) ", Sigma, " matrix")), side=1, line=3.5, at=0, adj=0.5, cex=1.75)

@@

#plot simulated correlations
heatmap.2(cor(t(expr)), scale = "none", trace = "none", col = colorpanel(50, "blue", "white", "red"),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
mtext(text = "(d) Simulated correlation", side=1, line=3.5, at=0, adj=0.5, cex=1.75)

#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "")
mtext(text = "samples", side=1, line=1.5, at=0.2, adj=0.5, cex=1.5)
mtext(text = "genes", side=4, line=1, at=-0.4, adj=0.5, cex=1.5)
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0, adj=0.5, cex=1.75)
```

```{r, include=FALSE}
pdf("Plotsimulation_smad-5.pdf", width = 13.44, height = 6.72)
#plot simulated expression data
heatmap.2(expr, scale = "none", trace = "none", col = bluered(50),
colsep = 1:length(V(graph)), rowsep = 1:length(V(graph)), labCol = "", mar = c(6, 6), keysize = 1)
mtext(text = "samples", side=1, line=1.5, at=0.55, adj=0.5, cex=1.5)
mtext(text = "genes", side=4, line=1, at=0.45, adj=0.5, cex=1.5)
mtext(text = "(e) Simulated expression data (log scale)", side=1, line=3.5, at=0.5, adj=0.5, cex=1.75)
dev.off()
system("sed -i.bak '/Plotsimulation_smad-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.md && rm paper.md.bak")
system("sed -i.bak '/Plotsimulation_smad-5/s/\\\\(.*\\\\)width=[.]415/\1width=.830/g' paper.tex && rm paper.tex.bak")
```
